---
title: "Hypothesis_Testing"
author: "Tejovardhan Medamarti"
date: "February 8, 2017"
<<<<<<< HEAD
output:
  pdf_document: default
  html_document: default
=======
output: html_document
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview and summary

Test hypotheses for the price of automobiles:

Source of the data can be found at : https://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data

Our objective is to find the hypothesis for the given below questions.

1. Compare and test Normality the distributions of price and log price - Use both a graphical method and a formal test.

2. Test significance of price (log price) stratified by a) fuel type, b) aspiration, and c) rear vs. front wheel drive. Use both graphical methods and the formal test.

3. Apply ANOVA to the auto price data to compare the price (or log price if closer to a Normal distribution) of autos stratified by number of doors, and body style - two sets of tests.

<<<<<<< HEAD
- Graphically explore the differences between the price conditioned by the categories of each variable -?Hint, make sure you have enough data for each category.
=======
- Graphically explore the differences between the price conditioned by the categories of each variable - Hint, make sure you have enough data for each category.
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92
- Use standard ANOVA and Tukey ANOVA to test the differences of these groups.

Note: Following packages are required to run the below report.

- dplyr


```{r packages, message=FALSE}
rm(list = ls())
require(dplyr)
require(ggplot2)

<<<<<<< HEAD
setwd("C:\\Tejo\\DataScience\\UW_Datascience_Course\\350\\DataScience350-master\\Lecture4\\Assignment")
=======
setwd("C:\\Tejo\\Datascience\\UW_Datascience_course\\350\\MethodsofDataAnalysis_350\\DataScience350-master\\Lecture4\\Assignment")

>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92
```

## Data loading and preparation

```{r dataloading , message=FALSE}
read.auto = function(file = 'Automobile price data _Raw_.csv'){
  ## Read the csv file
  auto.price <- read.csv(file, header = TRUE, 
                         stringsAsFactors = FALSE)
  
  ## Coerce some character columns to numeric
  numcols <- c('price', 'bore', 'stroke', 'horsepower', 'peak.rpm')
  auto.price[, numcols] <- lapply(auto.price[, numcols], as.numeric)
  
  ## Remove cases or rows with missing values. In this case we keep the 
  ## rows which do not have nas. 
  auto.price[complete.cases(auto.price), ]
}
auto.price = read.auto()

#Compare and test Normality the distributions of price and log price
#- Use both a graphical method and a formal test.

read.auto <- auto.price[auto.price$drive.wheels != "4wd",c("fuel.type", 
                                        "aspiration","drive.wheels" ,"price" ,
                                        "num.of.doors", "body.style" )]

read.auto <- read.auto[read.auto$num.of.doors != "?", ]
read.auto$log.price <- log(read.auto$price)
read.auto$scaled.log.price <- scale(read.auto$log.price, center = TRUE, scale = TRUE)
read.auto$scaled.price <- scale(read.auto$price, center = TRUE, scale = TRUE)

pop_auto.price = rnorm(nrow(read.auto), mean=mean(read.auto$scaled.price), sd = sd(read.auto$scaled.price))
pop_auto.norm = rnorm(nrow(read.auto), mean=0, sd = 1)
pop_auto.log.price = rnorm(nrow(read.auto), mean=mean(read.auto$scaled.log.price), sd = sd(read.auto$scaled.log.price))

```
**Lets start with the Question 1:**  

Compare and test Normality the distributions of price and log price - Use both a graphical method and a formal test.

**Null Hypothesis:** Distribution of log price data is identical to Standard Normal distribution


```{r kstestlog, message=FALSE}
ks.test(read.auto$scaled.log.price, pop_auto.norm)
```
<<<<<<< HEAD
- Based on the above results, we **ACCEPT** the null hypothesis as the P-values is > 0.05.
=======
- Based on the above results, we **REJECT** the null hypothesis as the P-values is < 0.5 (0.2352).
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

**Null Hypothesis:** Distribution of price data is identical to Standard Normal distribution

```{r kstestprice, message=FALSE}

ks.test(read.auto$scaled.price, pop_auto.norm)

```
<<<<<<< HEAD
**Conclusion:** Based on the above results, we **REJECT** the null hypothesis as the P-value is very small.
=======
**Conclusion:** Based on the above results, we **REJECT** the null hypothesis as the P-values is very small (0.001).
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

## Visualization:

Lets draw the QQ plot for the price and log price.

```{r visualization, message=FALSE}

par(mfrow = c(1, 2))

plot(sort(pop_auto.price), sort(pop_auto.norm), main = 'Plot of Price vs. normalized', 
     xlab = 'Sample Quantiles', ylab = 'Theoritical Quantiles')
abline(a = 0.0, b = 1.0, lty = 2, col = 'red')

plot(sort(pop_auto.log.price), sort(pop_auto.norm), main = 'Plot of log price vs. normalized', 
     xlab = 'Sample Quantiles', ylab = 'Theoritical Quantiles')
abline(a = 0.0, b = 1.0, lty = 2, col = 'red')

par(mfrow = c(1, 1))
<<<<<<< HEAD
=======

>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92
```

Lets draw a CDF plot for log price vs standard normalized data.

```{r standardization, message=FALSE}

# Have to standardize the x-values
x_seq = seq(-3,3,len=nrow(read.auto))
y_cdf1 = sapply(x_seq, function(x){
        sum(read.auto$scaled.log.price<x)/length(read.auto$scaled.log.price)
})
y_cdf2 = sapply(x_seq, function(x){
        sum(pop_auto.norm<x)/length(pop_auto.norm)
})

## Find the max deviation
k_s_stat = max(abs(y_cdf1 - y_cdf2))
k_s_stat
# where does it occur?
k_index = which.max(abs(y_cdf1-y_cdf2))
k_s_x = x_seq[k_index]
plot(x_seq,y_cdf1, col='blue', pch=16, main ='CDFs of standardized samples', 
     xlab = 'Value', ylab = 'Cumulative density')
points(x_seq,y_cdf2,col='red', pch=16) 
lines(c(k_s_x,k_s_x), c(y_cdf1[k_index],y_cdf2[k_index]),
      col='black', lwd=8)
```

Lets do our second question:

**2**. Test significance of price (log price) stratified by a) fuel type, b) aspiration, and c) rear vs. front wheel drive. Use both graphical methods and the formal test.


```{r plotfun, message=FALSE}

plot.t <- function(a, b, cols = c('pop_A', 'pop_B'), nbins = 20){
        maxs = max(c(max(a), max(b)))
        mins = min(c(min(a), min(b)))
        breaks = seq(maxs, mins, length.out = (nbins + 1))
        par(mfrow = c(2, 1))
        hist(a, breaks = breaks, main = paste('Histogram of', cols[1]), xlab = cols[1])
        abline(v = mean(a), lwd = 4, col = 'red')
        hist(b, breaks = breaks, main = paste('Histogram of', cols[2]), xlab = cols[2])
        abline(v = mean(b), lwd = 4, col = 'red')
        par(mfrow = c(1, 1))
}

```
For this analysis, i am stratifying the data by 10 records for each group.

<<<<<<< HEAD
**2(a). Lets start the analysis for log price by fuel type.**
=======
**1(a). Lets start the analysis for log price by fuel type.**
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

**Null Hypothesis:** Significance of log price by fuel type. There is no price difference with the fuel type.

```{r fueltype, message=FALSE}

n = 10

auto.fuel.type = read.auto %>% group_by(fuel.type ) %>% 
        sample_n(n, replace = FALSE)

auto.fuel.type.bydiesel <-  auto.fuel.type[auto.fuel.type$fuel.type=="diesel",]
auto.fuel.type.bygas <-  auto.fuel.type[auto.fuel.type$fuel.type=="gas",]

pop_A = auto.fuel.type.bydiesel$scaled.log.price
pop_B = auto.fuel.type.bygas$scaled.log.price

plot.t(pop_A, pop_B, cols = c("Log price by Diesel", "Log price by Gas"))

ks.test(pop_A, pop_B, alternative = "two.sided")

```
**Conclusion:**  Based on the above results, we failed to reject, hence we **ACCEPT** the null hypothesis.

<<<<<<< HEAD
**2(b). Lets start the analysis for log price by aspiration.**
=======
**1(b). Lets start the analysis for log price by aspiration.**
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

**Null Hypothesis:** Significance of log price by aspiration. There is no price difference with the aspiration.

```{r aspiration, message=FALSE}
auto.aspiration.type = read.auto %>% group_by(aspiration) %>% 
        sample_n(n, replace = FALSE)

auto.aspiration.type.std <-  auto.aspiration.type[auto.aspiration.type$aspiration=="std",]
auto.aspiration.type.turbo <-  auto.aspiration.type[auto.aspiration.type$aspiration=="turbo",]


pop_A = auto.aspiration.type.std$scaled.log.price
pop_B = auto.aspiration.type.turbo$scaled.log.price

plot.t(pop_A, pop_B, cols = c("Log price by std", "Log price by turbo"))

ks.test(pop_A, pop_B, alternative = "two.sided")

```
<<<<<<< HEAD
**Conclusion:** Based on the above results, we  **ACCEPT** the null hypothesis.

**2(C). Lets start the analysis for log price by drive wheels**
=======
**Conclusion:** Based on the above results, we  **REJECT** the null hypothesis.

**1(C). Lets start the analysis for log price by drive wheels**
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

**Null Hypothesis:** Significance of log price by drive wheels There is no price difference with the drive wheels.


```{r drivewheels, message=FALSE}

auto.drive.wheels = read.auto %>% group_by(drive.wheels) %>% 
        sample_n(n, replace = FALSE)

auto.drive.wheels.rwd <-  auto.drive.wheels[auto.drive.wheels$drive.wheels=="rwd",]
auto.drive.wheels.fwd <-  auto.drive.wheels[auto.drive.wheels$drive.wheels=="fwd",]


pop_A = auto.drive.wheels.rwd$scaled.log.price
pop_B = auto.drive.wheels.fwd$scaled.log.price
#plot.t(pop_A, pop_B)

ks.test(pop_A, pop_B, alternative = "two.sided")

plot.t(pop_A, pop_B, cols = c("Log price by rwd", "Log price by fwd"))

ks.test(pop_A, pop_B, alternative = "two.sided")
```
<<<<<<< HEAD
**Conclusion:** Based on the above results, we  **ACCEPT** the null hypothesis.
=======
**Conclusion:** Based on the above results, we  **REJECT** the null hypothesis.
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92

3. Apply ANOVA to the auto price data to compare the price (or log price if closer to a Normal distribution) of autos stratified by number of doors, and body style - two sets of tests.

**Null Hypothesis:** Significance of log price by Num of doors and body style. There is no price variances with the influence of Num of doors and body style.

```{r anova, message=FALSE}

auto.anova.sample = read.auto %>% group_by(num.of.doors, body.style) %>% 
        sample_n(n, replace = TRUE)

boxplot(auto.anova.sample$scaled.log.price ~ auto.anova.sample$num.of.doors+
                auto.anova.sample$body.style)


df_aov = aov(scaled.log.price ~ num.of.doors + body.style, data = auto.anova.sample)
summary(df_aov)
print(df_aov)

tukey_anova = TukeyHSD(df_aov)  # Tukey's Range test:
tukey_anova

plot(tukey_anova)

<<<<<<< HEAD
```

**Conclusion:**  Based on the above results,
1. "hatchback-convertible" body style has significant impact on the log price, hence we are **rejecting** the null hypothesis and for the rest of the other body styles, we are **accepting**.

2. For num.of.doors, we are **rejecting** the null hypothesis.

##Conclusion:

Log price is having signifant difference only when, 
- body style equals "hatch back convertible" 
- the num of doors.


=======
library(agricolae)
HSD.test(ajuste, 'chocolate$Tipo')

```
>>>>>>> bac3d7b3bc81153dc189c72ca22c42969e9bda92
