---
title: "Analysis of switching strategy for Monty Hall problem"
author: "Tejovardhan Medamarti"
date: "31 January 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview and summary
The Monty Hall problem is a brain teaser, in the form of a probability puzzle (Gruber, Krauss and others), loosely based on the American television game show **Let's Make a Deal** and named after its original host, Monty Hall. The problem was originally posed in a letter by Steve Selvin to the American Statistician in 1975 (Selvin 1975a), (Selvin 1975b).

Our objective in this assignment is to write an R-script to simulate the conditional probabilities and variances of the distributions for winning a car vs. a goat for the Monty Hall Problem. Include the cases of switching or not switching. Compare these cases to determine the best strategy.

Note: Following packages are required to run the below report.
-plyr
-ggplot2
-scales

```{r loadlibrary}
rm(list=ls())
require (plyr)
require(ggplot2)
require(scales)
```

## Data Preparation:

```{r preparedata}
MontyHall.Problem.Simulation <- function (noofgates, r, strategy)
{
  dfmontymatrix <- data.frame(matrix(0 ,ncol = 6, nrow = r))
  cnames <- c("Gatechosen", "PrizeGate","MontySelection","Strategy", "Result", "Noofsimulations")
  colnames(dfmontymatrix) <- cnames
  #noofgates <- 1:3

  dfmontymatrix[, "Noofsimulations"] <- r  # No of simulations.
  for (ri in 1: r)
  {
    indexprize <- sample(noofgates, 1, replace = TRUE) 
    Gatechosen <- sample(noofgates, 1, replace = TRUE)
    
    if(indexprize != Gatechosen)
      {
      MontySelection <- noofgates[-c(indexprize, Gatechosen)]
    } else 
      { 
          MontySelection <- sample(noofgates[-c(indexprize, Gatechosen)], 1)
      } 
    
    dfmontymatrix[ri, "Gatechosen"] <- Gatechosen
    dfmontymatrix[ri, "PrizeGate"] <- indexprize
    dfmontymatrix[ri, "MontySelection"] <- MontySelection
    dfmontymatrix[ri, "Strategy"] <- sample(strategy, 1, replace = TRUE) 
   
    if (dfmontymatrix[ri,"Strategy"] == "stay")
    {
      if (dfmontymatrix[ri,"Gatechosen"] == dfmontymatrix[ri, "PrizeGate"])
      {
        dfmontymatrix[ri, "Result"] = "win"
      } 
      else 
      {
        dfmontymatrix[ri, "Result"] = "loss"
      } 
    } else
    {
      
      if (dfmontymatrix[ri,"Gatechosen"] != dfmontymatrix[ri, "PrizeGate"])
      {
        dfmontymatrix[ri, "Result"] = "win"
      } 
      else 
      {
        dfmontymatrix[ri, "Result"] = "loss"
      }   
    }
  }
  
  return (with (dfmontymatrix, table(Noofsimulations, Strategy, Result)))
  
}
```
## Data simulation:
Lets simulate the data with datapoints (1000, 10000, 20000)

```{r datasimulation}

r <- c( 1000, 10000, 20000)
noofgates <-1:3
strategy <- c("stay", "switch") 
sampledata <- lapply(r, function(r) MontyHall.Problem.Simulation(noofgates, r, strategy))
final.output <- ldply (sampledata, data.frame)
final.output$Noofsimulations <- as.integer(as.character(final.output$Noofsimulations))
final.output <- mutate(final.output, 
                       Percentage = percent(Freq/ Noofsimulations))

print (final.output)

```
## Visualization:
```{r visualization}

ggplot(final.output, aes(Result,Freq, fill = Strategy ))+
  geom_bar( stat = "identity")+
  geom_text(aes(label = Percentage))+
  facet_wrap(~Noofsimulations, scales = "free" )+
  labs(title = "Analysis of strategies involved for Monty Hall problem ", y = "No of Simulations", x = "Strategy Result")

```

**Lets do only for strategy = "switch"**

```{r visualization 1}

strategy <-  "switch" 

sampledata <- lapply(r, function(r) MontyHall.Problem.Simulation(noofgates, r, strategy))

final.output <- ldply (sampledata, data.frame)
final.output$Noofsimulations <- as.integer(as.character(final.output$Noofsimulations))
final.output <- mutate(final.output, 
                       Percentage = percent(Freq/ Noofsimulations))

ggplot(final.output, aes(Result,Freq, fill = Strategy ))+
  geom_bar( stat = "identity")+
  geom_text(aes(label = Percentage))+
  facet_wrap(~Noofsimulations, scales = "free" )+
  labs(title = "Analysis of switching strategy for Monty Hall problem ", y = "No of Simulations", x = "Strategy Result")

```

**Lets do only for strategy = "stay"**

```{r visualization 2}
strategy <-  "stay" 

sampledata <- lapply(r, function(r) MontyHall.Problem.Simulation(noofgates, r, strategy))

final.output <- ldply (sampledata, data.frame)
final.output$Noofsimulations <- as.integer(as.character(final.output$Noofsimulations))
final.output <- mutate(final.output, 
                       Percentage = percent(Freq/ Noofsimulations))

ggplot(final.output, aes(Result,Freq, fill = Strategy ))+
  geom_bar( stat = "identity")+
  geom_text(aes(label = Percentage))+
  facet_wrap(~Noofsimulations, scales = "free" )+
  labs(title = "Analysis of not switching strategy for Monty Hall problem ", y = "No of Simulations", x = "Strategy Result")

```

##Conclusion:

Based on the above visualizations, Switching strategy has more chances to win than staying.
