---
title: "BayesianTest"
author: "Tejovardhan Medamarti"
date: "February 21, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview and summary

Probability of texting:
  
 - You are asked to compute the probability that the driver of a car is texting at a specific intersection.
 
 - Nationally the cumulative probability that a driver is texting is:
 
   **P = 0.5, at x = 0.1**
   
   **P = 0.75 at x = 0.3**
   
 - You observe cars at a location three times and note the number of texting drivers:
 
- 2 texting out of 20 drivers
- 4 texting out of 20 drivers
- 1 texting out of 20 drivers

-- Given these data

- Compute the Beta prior, and report the coefficients

 - Plot the prior, likelihood and posterior three times as you update your belief based on collecting more data
 
- Simulate the final posterior distribution and do the following:

  1. Plot the posterior with the 90% HDI shown
  2. Report the upper and lower limits of the 90% HDI
  3. Of the next hundred drivers what are the number of texting drivers in the 90% HDI?
  4. Are the drivers in this area better or worse that the national figures indicate?
  
  **Note**: Following packages are required to run the below report.

- LearnBayes
- Knitr

```{r bayesian, message=FALSE, echo = FALSE}
require(LearnBayes)
```
### Lets create a beta  distribution, named beta.par.

```{r betapar}
beta.par <- beta.select(list(p=0.5, x=0.1), list(p=0.75, x=.3))

```
##Q1)

- 2 texting out of 20 drivers
- 4 texting out of 20 drivers
- 1 texting out of 20 drivers

- Compute the Beta prior, and report the coefficients

 - Plot the prior, likelihood and posterior three times as you update your belief based on collecting more data
 

```{r plotthree, fig.width=6, fig.height=12}

par(mfrow = c(3,1))
beta.par + c(2, 18)
triplot(beta.par, c(2, 18))

beta.par + c(4, 16)
triplot(beta.par, c(4, 16))

beta.par + c(1, 19)
triplot(beta.par, c(1, 19))
par(mfrow = c(1,1))

```

##Q2)

Simulate the final posterior distribution and do the following:

  1. Plot the posterior with the 90% HDI shown
  2. Report the upper and lower limits of the 90% HDI
  3. Of the next hundred drivers what are the number of texting drivers in the 90% HDI?

```{r simulation}
options(repr.plot.width=8, repr.plot.height=5)
beta.post.par <- beta.par + c(7, 53)
post.sample <- rbeta(100, beta.post.par[1], beta.post.par[2])
par(mfrow = c(1,2))
quants = quantile(post.sample, c(0.05, 0.95))
breaks = seq(min(post.sample), max(post.sample), length.out = 41)
hist(post.sample, breaks = breaks, 
     main = 'Distribution of samples \n with 90% HDI',
     xlab = 'Sample value',
     ylab = 'Density')
abline(v = quants[1], lty = 3, col = 'red', lwd = 3)
abline(v = quants[2], lty = 3, col = 'red', lwd = 3)
qqnorm(post.sample)
par(mfrow = c(1,1))

```

```{r predplot}

predplot(beta.post.par, 7, 53)

n <- 100
s <- 0:n
pred.probs <- pbetap(beta.post.par, n, s)
pred.probs.beta <- pbetap(beta.par, n, s)

plot(s, pred.probs, type="h", 
     main = paste('Probability distribution of successes in', as.character(n), 'trials'),
     xlab = 'Successes')

discint(cbind(s, pred.probs.beta) , 0.90)
discint(cbind(s, pred.probs), 0.90)


```
### Visualization: 

```{r visn, fig.width=6, fig.height=12}
par(mfrow = c(2,1))


plot(s, pred.probs.beta, type="h", 
     main = paste('Probability distribution of Prior beleifs in', as.character(n), 'trials'),
     xlab = 'Successes')

plot(s, pred.probs, type="h", 
     main = paste('Probability distribution of posterier beleifs in', as.character(n), 'trials'),
     xlab = 'Successes')

par(mfrow = c(1,1))

```

###Conclusion: 
Based on the above results and visualizations, Cumulative probability of the city is better than the Nationally cumulative probability that a driver is texting when driving.
