---
title: "Time Series Analysis"
author: "Tejovardhan Medamarti"
date: "March 7, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview and summary

Our objective in this assignment is to perform time series analysis on the data for one of Milk Production, or Ice Cream Production, in the CADairyProduction.csv.

We have chosen to analyse the data for the ice cream Production.

file to answer the following questions

  - Is this time series stationary?
  
  - Is there a significant seasonal component?
  
  - For the residual from the STL decomposition of the time series what is the order of the ARMA(p,q) process that best fits?
  
  - Forecast production for 12 months. Report both numeric values and plot the confidence intervals. Are the confidence intervals reasonably small compared to the forecast means? How do the confidence intervals behave as time moves to the future?
  
**Note: ** Following packages are required to run the below packages.
  - forecast
  - repr
```{r pkgrequired, warning=FALSE, message=FALSE}
require(forecast)
require(repr)

```


### Data staging:

```{r CAdairystaging, warning=FALSE, message=FALSE}
read.CADairy = function(file = "C:\\Tejo\\Datascience\\350\\Classwork\\DSClasswork\\Lecture8\\Assignment\\CADairyProduction.CSV"){
    ## Read the csv file
    dairy.produce <- read.csv(file, header = TRUE, 
                              stringsAsFactors = FALSE)
    
  }
  
ca.dairy = read.CADairy()

```
Since we are dealing only with the Icecream prod feature, lets create a basic time series vector vec.ts which only contains Icecream prod feature.

```{r}
vec.ts = with(ca.dairy, log(ts(Icecream.Prod, start = c(1995, 1), freq = 12)))
attributes(vec.ts) # Note the time series attributes

options(repr.pmales.extlot.width=8, repr.plot.height=4)
plot(vec.ts) # Note the x-axis is the time attribute

```

**Lets create a helper functions for the visualization purpose.**

```{r utilfun, warning=FALSE, message=FALSE}

plot.acf <- function(df, col = 'remainder', is.df =TRUE){
  if(is.df) temp <- df[, col]
  else temp <- df
  par(mfrow = c(2,1))
  acf(temp, main = paste('ACF of', col))
  pacf(temp, main = paste('PACF of', col))
  par(mfrow = c(1,1))
}

## Function for ARIMA model estimation
ts.model = function(ts, col = 'remainder', order = c(0,0,1)){
  mod = arima(ts, order = order, include.mean = FALSE)
 print(mod)
  mod
}


```


Q1: Is this time series stationary?

#### - Plot 1:

```{r plot1, echo=FALSE}
plot.acf(vec.ts, col = 'time series - Is stationary', is.df = F)
```

```{r ljung}
Box.test(vec.ts, lag =20, type="Ljung-Box")
```

**Result:** Based on the above results, ACF and PACF are dying slowly and also the LJung-box test reveals the ts is stationary based on the p-value.Hence the **time series is stationary**.

Q2: Is there a significant seasonal component??
#### - Plot 2:

```{r tssignificant}
plot(vec.ts, main = 'seasonal time series')
monthplot(vec.ts)
```

**Result:** Based on the above results, **time series is a significant seasonal component**.

Q3: For the residual from the STL decomposition of the time series what is the order of the ARMA(p,q) process that best fits?

```{r tsdecomp}
## Decomposition of the time series into components
ts.decomp <- function(df, col = 'dairy.ts', span = 0.25, Mult = TRUE, is.df = TRUE){
  # if(Mult) temp = log(df[, col])  else temp = ts(df[, col]
  if(is.df) temp = log(df[, col])  
  else temp = df
  spans = span * length(temp)  
  fit <- stl(temp, s.window = "periodic", t.window = spans)
  plot(fit, main = paste('Decompositon of',col,'with lowess span = ', as.character(span)))
  fit$time.series
}

temp = ts.decomp(vec.ts, is.df = FALSE, Mult = FALSE)

```
#### - Plot 3:
```{r plot3}
plot.acf(temp[,3],is.df=FALSE)
iceCream.arima = ts.model(temp[, 3], col = 'ARIMA model for icecream production', order = c(2,0,2))##
plot.acf(iceCream.arima$resid[-1],is.df=FALSE)

```

**Result:** Based on the above results, we are getting best fit at order c(2,0,2) for the ARMA(p,q) process.

**Q4:** Forecast production for 12 months. Report both numeric values and plot the confidence intervals.

  - Are the confidence intervals reasonably small compared to the forecast means? 
  
  - How do the confidence intervals behave as time moves to the future?
  

```{r tsforecast}

fit.icecream = auto.arima(vec.ts, max.p=3, max.q=3,
                       max.P=2, max.Q=2, max.order=5, max.d=2, max.D=1,
                       start.p=0, start.q=0, start.P=0, start.Q=0)
summary(fit.icecream)
icecream.forecast = forecast(fit.icecream, h=12)
summary(icecream.forecast)
plot(icecream.forecast)

```

####Conclusion:
  Based on the above results, Confidence intervals are relatively small when compared to the forecasted means and they behave as the time moves.

