---
title: "Lesson1_Assignment_EnergyEfficiency"
author: "Suman Nooney"
date: "January 18, 2017"
output:
  pdf_document: default
  html_document: default
email: "suman.nooney@gmail.com"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Synopsis

This report is based on the dataset available at <https://archive.ics.uci.edu/ml/datasets/Energy+efficiency#>. Our interest in this data, is 
to find out the 3 interesting patterns how the Heating and Cooling Loads are impacted with the given 8 input variables.

### Variable(s) Information:

* Relative Compactness
* Surface Area - m²
* Wall Area - m²
* Roof Area - m²
* Overall Height - m
* Orientation - 2:North, 3:East, 4:South, 5:West
* Glazing Area - 0%, 10%, 25%, 40% (of floor area)
* Glazing Area Distribution (Variance) - 1:Uniform, 2:North, 3:East, 4:South, 5:West
* Heating Load - kWh/m²
* Cooling Load - kWh/m²

## Data staging
There is a cleansed dataset available from the Github at <https://raw.githubusercontent.com/StephenElston/DataScience350/master/Lecture1/EnergyEfficiencyData.csv>.  Used it as a source, and below are the steps used in downloading and created a categorical variables for **"Orientation"**, **"Glazing Area Distribution (variance)"**.

```{r}
rm(list = ls())

SourceURL_Raw <- "https://raw.githubusercontent.com/StephenElston/DataScience350/master/Lecture1/EnergyEfficiencyData.csv"

energy.efficiency <- read.csv( SourceURL_Raw, header = TRUE)

require(ggplot2)
#install.packages("gridExtra")
require(gridExtra)

energy.efficiency$Orientation <- as.factor(energy.efficiency$Orientation)

levels(energy.efficiency$Orientation) <- c("North", "East", "South", "West")

energy.efficiency$Glazing.Area.Distribution <- as.factor(energy.efficiency$Glazing.Area.Distribution)

levels(energy.efficiency$Glazing.Area.Distribution) <- c("UnKnown", "Uniform", "North", "East", "South", "West")

energy.efficiency$Glazing.Area <- as.factor(energy.efficiency$Glazing.Area)

levels(energy.efficiency$Glazing.Area) <- c("0%", "10%", "25%", "40%")
```
Lets look at the summary of the energy.effiiency data.
```{r}
summary(energy.efficiency)
```

## Visualizations
### Plot 1:
Lets visualize how the overall height impacts the overall Cooling and Heating Load using density plot.
        
```{r }
ggplot(energy.efficiency, aes(x = Heating.Load , y = Cooling.Load)) + 
        geom_point( aes(col = factor(Overall.Height)), alpha= 0.3) +
        geom_density2d()+
        xlab('Heating Load') + 
        ylab('Cooling Load') + 
        ggtitle('Heat and Cold Load Comparison by Overall Height')

```

From the above plot, it is clear that the overall height plays a critical role in heating and cooling load.

### Plot 2:
Lets visualize our second plot by how the roof and wall areas impacts the Heating Load using box plot.
```{r echo=TRUE}
ggplot(energy.efficiency, aes(x = factor(Roof.Area), y = Heating.Load,  group = Surface.Area)) +
        geom_boxplot(aes(fill = factor(Wall.Area))) + 
        # geom_jitter(alpha= 0.3)+
        facet_grid(.~Overall.Height)+
        xlab('Roof Area') + 
        ylab('Heating Load') + 
        ggtitle('Distribution of Heating Load on Roof Area by Wall Area and Overall Height')
```

###Plot: 3
Lets visualize the load distribution of Glazing Area by Orientation.
        
```{r}
ggplot(energy.efficiency, aes(Heating.Load, Cooling.Load))+
  geom_point(aes(col = factor(Overall.Height)))+
  facet_grid(Orientation ~ Glazing.Area)+
  xlab('Heating Load') + 
  ylab('Cooling Load') + 
  ggtitle(' Heating and Cooling load distribution by Orientation and Glazing Area.')

```


###Plot: 4
Lets visualize the Cooling and Heating load distribution by Orientation and Roof Area.

```{r}
ggplot(energy.efficiency, aes(x = Cooling.Load, y = Heating.Load))+
        geom_point(aes(colour= Orientation))+
        facet_grid(Overall.Height ~ Roof.Area) 

```

### Conclusion:
From the above plots, we have clearly observed the Overall Height has a significant impact on overall heating and cooling load.
